---
title: "Mod2_LinearRegression"
author: "Grace Williams"
date: "2023-05-24"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Simple Linear Regression and Correlation

#### Libraries and Datasets Needed
```{r}
library(tidyverse)
library(tidymodels)
library(GGally)
library(lmtest)

air = airquality
```

### Question 1: How many rows are in the "air" dataframe?
##**153**
```{r}
nrow(air)
```

### Question 2: How many columns are in the "air" datagrame?
##**6**
```{r}
ncol(air)
```

### Question 3: True/False: There is missing data in "Ozone" variable in the dataframe.
##**TRUE**
```{r}
summary(air$Ozone)

is.na(air$Ozone)
```

### Question 4: Which variable is the most likely to be the response (Y) variable?
### A. Ozone
### B. Solar.R
### C. Wind
### D. Temp
### E. Month
### F. Day
##**D. Temp**

### Cleaning up the data.
```{r}
air2 = air %>% drop_na()
```

### Question 5: How many rows in this new (air2) data frame?
##**111**
```{r}
nrow(air2)
ncol(air2)
```

### Question 6: How many columns remain in this new (air2) data frame?
##**6**


### Use the “ggpairs” function to develop a visualization of the relationships in this dataset and to show correlation values for the combinations of variables. Then use the “ggcorr” function to develop a correlation matrix for the variables. Hint: Use “label = TRUE” in the “ggcorr” function to show the correlation values.

### Question 7: Which variable is the most strongly correlated with the "Ozone" variable?
### A. Solar.R
### B. Wind
### C. Temp
### D. Month
### E. Day
##**C. Temp**
```{r}
ggpairs(air2)
ggcorr(air2, label = TRUE)
```
### Question 8: Which variable is least strongly correlated with the "Ozone" variable?
### A. Solar.R
### B. Wind
### C. Temp
### D. Month
### E. Day
##**E. Day**

### Question 9: Plot "Temp" (x axis) versus "Ozone (y axis) using the "ggplot" function. Choose an appropriate chart type. Which statement best describes the relationship between "Temp" and "Ozone"?
### A. As Temp increases, Ozone decreases
### B. As Temp increases there is no noticeable change in Ozone
### C. As Temp increases, Ozone increase
##**C. As Temp increases, Ozone increases**
```{r}
ggplot(air2, aes(x = Temp, y = Ozone))+
  geom_point()+
  geom_smooth(method=lm)
```

### Use Tidymodels to create a linear regression model using “Temp” to predict “Ozone”. You miss wish to callyour model fit “lm_fit”.
```{r}

OT2 = recipe(Ozone ~ Temp, air2)
OT2

lm_model = 
  linear_reg() %>%
  set_engine("lm")
lm_model

lm_workflow =
  workflow() %>%
  add_model(lm_model) %>%
  add_recipe(OT2)
lm_workflow

lm_fit = fit(lm_workflow, air2)

summary(lm_fit$fit$fit$fit)

confint(lm_fit$fit$fit$fit)


testdata = data.frame(Temp = 80)
predict(lm_fit, testdata)

```
### Question 10: What is the slope of this regression model (to four decimal places)?
##**2.4391**

### Question 11: What is the R-squared value of this model (not Adjusted R-squared) (to three decimal places)?
##**0.488**

### Question 12: Is the “Temp” variables significant in the model?
##**Yes, Temp is significant**

### Use the code below to generate 95% confidence intervals for the coefficients. Note that you may need to change “lm_fit” to the name of your model fit if you used a different name. 
#### confint(lm_fit$fit$fit$fit)

### Question 13: True/False: A 95% confidence interval for the slope coefficient does not contain zero.
##**TRUE**

### Question 14:  Using your linear regression model with “Temp” to predict “Ozone”, what is the predicted “Ozone” value when “Temp” is equal to 80 (to two decimal places)?
##**47.5**

### Perform appropriate model diagnostics to verify whether or not the model appears to meet the four linear regression model assumptions.

### Question 15: True/False: There is no evidence of non-independent (autocorrelated) residuals.
##**True**
```{r}
hist(air2$Temp)

plot(Ozone ~ Temp, air2)

cor(air2$Ozone, air2$Temp)
```

