---
title: "Mod6_Assign1-Quiz"
author: "Grace Williams"
date: "2023-06-22"
output: word_document
editor_options: 
  chunk_output_type: console
---
```{r}
library(tidyverse)
library(tidymodels)
library(readr)
library(cluster)
library(factoextra)

trucks <- read_csv("trucks-1.csv")
```

```{r}
na.omit(trucks)
```

```{r}
summary(trucks)
trucks_scaled = scale(trucks)
summary(trucks_scaled)
```
 
```{r}

set.seed(1234)
clusts = 
  tibble(k = 1:10) %>% #try from 1 to 10 clusters
  mutate(
    kclust = map(k, ~kmeans(trucks_scaled, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_scaled)
  )

clusts
```

```{r}
clusters =
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>%
  unnest(cols = c(augmented))

clusterings =
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```

```{r}
cust_clust = kmeans(trucks_scaled, centers = 4)
cust_clust
```

```{r}
trucks <- augment(cust_clust, trucks)

head(trucks)
```

```{r}
ggplot(trucks, aes(x=Distance,y=Speeding,color=factor(.cluster))) + geom_point()
```

# Questions:

# 1. Which characteristics (select all that apply) of the relationship between Distance and Speeding seem most apparent?
## *a. There appears to be more speeding among the drivers with smaller Distances 
## *b. The data points are arranged in what appear to be four clusters 
## *c. Longer distance drivers appear more likely to speed 
## *d. There are no well-defined clusters of data points
### b and c

# 2. What is the maximum value (to four decimal places) of the Distance variable in the scaled dataset?
### 3.1559

# 3. Which statement best describes the resulting clusters?
## *a. Drivers with shorter distances are in one cluster and those with longer distances are in another 
## *b. Drivers with a higher proportion of speeding are in one cluster and those with a lower proportion of speeding are in another 
## *c. Neither of these statements apply to the resulting clusters 
### a

## 4. Which value of k appears to be most appropriate for this data?
### 4

## 5. What number of clusters appears to be ideal based on this plot?
### 4

## 6. Which statements (select all that apply) appear to be most apparent about the clusters created in this question?
## *a. One cluster is composed of short distance drivers with a low proportion of speeding. 
## *b. One cluster is composed of long distance drivers with a high proportion of speeding. 
## *c. One cluster is composed of long distance drivers with a low proportion of speeding. 

## *d. One cluster is composed of short distance drivers with a high proportion of speeding. 
### All of the above